Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-01-16 paper/plos.tex
edited intro to reference thicker silicon slits and added some references Tested everything on crystal.

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria, mannik2009bacterial}. They have
experimentally forced \emph{E. coli} cells into microfabricated
silicon channels of $0.25\micron$ thickness. Upon entering the
channels, the cells undergo a mechanical deformation in which they
both widen and lengthen within the plane of the channel.  This
deformation results in very wide (they can reach widths of over
$5\micron$) flattened cells that when viewed from the top down have
irregular and asymmetric shapes.  While these cells are still able to
divide into surprisingly equal volumes, the MinD oscillations in these
cells are irregular, both temporally and spatially. Seen with
fluorescent microscopy, the MinD maximize in multiple locations within
the cell, in a seemingly random sequence. These experiments allow for
an opportunity to test MinD simulation models against more extreme
cases than have been seen thus far.

A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings
published after his 2003 paper \cite{Meacci2007,Loose2011}. These
models introduce slow diffusion along the membrane (two to three
orders of magnitude slower than that within the cytoplasm), revise the
process of MinD recruitment to the membrane, and change some of the
constants used in the model. The Bonny 2013 model assumes an
instantaneous nucleotide exchange rate, and allows MinE to remain
independently attached to the membrane after dissasociating with
MinD~\cite{bonny2013membrane}.
%

@@

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic} \fixme{search for more}, in both
deterministic and stochastic variants to study flattened cells similar
to those observed by Mannik \emph{et al.}~\cite{mannik2009bacterial}.
\fixme{Some weak justification... it has been widely used.}

>>>>

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2012robustness,mannik2009bacterial,mannik2010bacteria}. They
have experimentally forced \emph{E. coli} cells into microfabricated
channels of $0.25\micron$ thickness, in which they area able to grow
and divide. The channels, which are covered in a layer of PDMS, are
deformable to some degree.  Experiments with a different geometry that
is constructed without PDMS show that it's reasonable to assume the
bacteria deform the PDMS channels to a thickness of at least
$0.4\micron$~\cite{mannik2009bacterial,mannik2010bacteria}.

Upon entering the channels, the cells undergo a mechanical deformation
in which they both widen and lengthen within the plane of the channel.
This deformation results in very wide (they can reach widths of over
$5\micron$) flattened cells that when viewed from the top down have
irregular and asymmetric shapes.  While these cells are still able to
divide into surprisingly equal volumes, the MinD oscillations in these
cells are irregular, both temporally and spatially. Seen with
fluorescent microscopy, the MinD maximize in multiple locations within
the cell, in a seemingly random sequence. These experiments allow for
an opportunity to test MinD simulation models against more extreme
cases than have been seen thus far.

A number of models of the MinD protein system have been developed that
accurately describe its basic oscillatory nature.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
approach with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachment of MinD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.
%
Several models
\cite{fange2006noise,bonny2013membrane,halatek2012highly} modify
Huang's original model to more accurately treat experimental findings
published after his 2003 paper
\cite{meacci2006mobility,loose2011min}. These models introduce slow
diffusion along the membrane (two to three orders of magnitude slower
than that within the cytoplasm), revise the process of MinD
recruitment to the membrane, and change some of the constants used in
the model. The Bonny 2013 model assumes an instantaneous nucleotide
exchange rate, and allows MinE to remain independently attached to the
membrane after dissasociating with MinD~\cite{bonny2013membrane}.
%

@@

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic} \fixme{search for more} in both
deterministic and stochastic variants, to study flattened cells of
$0.4\micron$ thickness that are similar to those observed by Mannik
\emph{et al.}~\cite{mannik2009bacterial}.  \fixme{Some weak
  justification... it has been widely used.}
